---
title: "Batting"
author: "Jim Albert"
date: "2022-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

### Illustration of some tidyverse work to explore the Batting table in the Lahman database

I load in three packages:

- the Lahman package contains the baseball data
- the dplyr package is part of the tidyverse suite of packages useful for manipulation of data frames
- the ggplot2 package allows us to create graphs

```{r}
library(Lahman)
library(dplyr)
library(ggplot2)
```

The Lahman database contains season-to-season data for MLB seasons from the beginning to the 2021 season.

The Batting table is a data frame with 110,495 observations with 22 variables.

**Question:  Which player had the highest career batting average between 1990 and 2020?**

Here I start with the Batting data frame:

-- filter the data to get only the stats for the seasons between 1990 and 2020
-- for each player, find the cumulative H and AB (do this by the group_by and summarize functions)

```{r}
Batting %>% 
  filter(yearID >= 1990,
         yearID <= 2020) %>% 
  group_by(playerID) %>% 
  summarize(H = sum(H),
            AB = sum(AB)) -> S
```

Here is what this new data frame S looks like:

```{r}
head(S)
```

I only want to consider players who have had at least 1000 AB during this period.  I will call the new data frame S1000.

```{r}
S %>% 
  filter(AB >= 1000) -> S1000
```

Next, I compute the batting average for each player -- round to three decimal places.

```{r}
S1000 %>% 
  mutate(AVG = round(H / AB, 3)) -> S1000
```

Last, I sort this data frame by AVG from highest to lowest:

```{r}
S1000 %>% 
  arrange(desc(AVG)) -> S1000
```

Use the head function to show the top 10:

```{r}
head(S1000, 10) -> Top_AVG
Top_AVG
```

It would be helpful to add the names of these players.  The People data frame contains info about each player including their player code, first and last names.  I will select specific variables from the People data frame and merge this information with my batting tab.e

```{r}
inner_join(Top_AVG, 
           select(People, playerID,
                  nameFirst, nameLast),
           by = "playerID") -> Top_AVG
```

Now I display the names and top averages:

```{r}
Top_AVG %>% 
  select(nameFirst, nameLast, AVG)
```

I create a new variable Name that pastes the first and last names.  

```{r}
Top_AVG %>% 
  mutate(Name = paste(nameFirst, nameLast)) ->
  Top_AVG
```

Then I use a dotplot to display these top averages.

The ggplot() function has these inputs:

- first I indicate the data frame to be used
- the aes() function indicates roles -- here I want the x variable to be AVG and the y variable to be Name
- the geom_point() is a geometric object to display - here this is a point.
- I add a title to indicate what I am graphing

```{r}
ggplot(Top_AVG,
       aes(AVG, Name)) +
  geom_point() +
  labs(title = "Top Batting Averages in Period 1990:2020")
```

Clearly, Tony Gwynn was the best hitter for average during this period.