---
title: "November 10"
author: "Jim Albert"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

### Data frame manipulations

In the `dplyr` package, here are some verbs for data manipulation:

- `select` will choose variables (columns)
- `filter` will choose observations (rows)
- `mutate` will define new variables
- `arrange` will sort the data frame with respect to somoe variable
- `summarize` will compute a data summary 

The piping symbol `%>%` passes the data frame from one operation to another operation.

A common thing to do is to divide the data frame by a `group_by` on some variable and `summarize` each subgroup.

### Exploring Lahman's Pitching Table

Suppose we are interested in finding the pitchers with the highest strikeout rates (9 x SO / IP) in the seasons 1980 through 2021.  Interested in 

- the highest rates for a particular season 
- the highest career rates

Load the `Lahman` and `dplyr` packages.

```{r}
library(Lahman)
library(dplyr)
```

Start with Pitching data frame.

- filter to select the rows with the right seasons
- for each player/season, compute the number of SO and IP
- store results in data frame S

```{r}
Pitching %>% 
  filter(yearID >= 1980, yearID <= 2021) %>% 
  group_by(playerID, yearID) %>%
  summarize(SO = sum(SO),
            IP = sum(IPouts / 3)) -> S
head(S)
```

Focus on pitchers who threw at least 100 innings in a season.

```{r}
S %>% 
  filter(IP >= 100) -> S100
```

Compute the SO rates and sort table from high to low.
Display the top 10.

```{r}
S100 %>% 
  mutate(SO_Rate = 9 * SO / IP) %>% 
  arrange(desc(SO_Rate)) -> S100
head(S100)
```

Attach the names to table.

```{r}
inner_join(select(People, playerID,
                  nameFirst, nameLast),
           S100, by = "playerID") %>% 
  mutate(Name = paste(nameFirst, nameLast)) %>% 
  select(Name, yearID, SO_Rate)  %>% 
  arrange(desc(SO_Rate)) -> S100
head(S100)
```

Might be interesting to look at bottom 10.

```{r}
tail(S100)
```

Now we look at cumulative strikeout rates over this period (min IP is 1000).

Here we are grouping by pitcher (ignoring season) and finding total SO and IP for each pitcher.

```{r}
Pitching %>% 
  filter(yearID >= 1980, yearID <= 2021) %>% 
  group_by(playerID) %>%
  summarize(SO = sum(SO),
            IP = sum(IPouts / 3)) %>% 
  filter(IP >= 1000) %>% 
  mutate(SO_Rate = 9 * SO / IP) -> S_cum 
```

Add the name information from the People data frame.

```{r}
  inner_join(select(People, playerID,
                  nameFirst, nameLast),
             S_cum,
             by = "playerID") -> S_cum 
```

Create a top 6 table.

```{r}
S_cum %>% 
  mutate(Name = paste(nameFirst, nameLast)) %>% 
  select(Name, IP, SO_Rate) %>% 
  arrange(desc(SO_Rate)) %>% 
  head(6)
```

### Retrosheet Data

Retrosheet provides play by play data for all plays in recent seasons.

I have put the Retrosheet file for a single season on my Github respository site.

I use the read_csv() function from the readr package to read in this data and store in the data frame d2021.

```{r}
library(readr)
```

```{r}
d2021 <- read_csv("https://raw.githubusercontent.com/bayesball/HomeRuns2021/main/retro2021.csv")
```
Show one line from this file.

```{r}
d2021[10000, ]
```


