---
title: "World Series"
author: "Jim Albert"
date: "2022-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

### Illustration of some tidyverse work to explore the SeriesPost table in the Lahman database

I load in three packages:

- the Lahman package contains the baseball data
- the dplyr package is part of the tidyverse suite of packages useful for manipulation of data frames
- the ggplot2 package allows us to create graphs

```{r}
library(Lahman)
library(dplyr)
library(ggplot2)
```

The Lahman database contains season-to-season data for MLB seasons from the beginning to the 2021 season.

The SeriesPost table is a data frame with 367 observations with 9 variables.  This data frame gives information about every post season MLB series including the World Series.

**Question:  Which teams have played in the World Series the most?**

Here I start with the SeriesPost data frame:

-- filter the data to only consider the World Series playoffs
-- I will separately look at the WS winners and losers

```{r}
SeriesPost %>% 
  filter(round == "WS") -> WorldSeries
head(WorldSeries)
```

```{r}
WorldSeries %>% 
  group_by(teamIDwinner) %>% 
  summarize(N = n()) %>% 
  arrange(desc(N)) -> Winners
head(Winners)
```

```{r}
WorldSeries %>% 
  group_by(teamIDloser) %>% 
  summarize(N = n()) %>% 
  arrange(desc(N)) -> Losers
head(Losers)
```

Merge the two data frames.

```{r}
inner_join(Winners, Losers,
           by = c("teamIDwinner" = "teamIDloser")) ->
  WL
head(WL)
```

Add the number of times winner to the number of times loser and rank from high to low.

```{r}
WL %>% 
  mutate(Times = N.x + N.y) %>% 
  arrange(desc(Times)) %>% 
  head(10)
```

Which team is NY1?

```{r}
Teams %>% 
  filter(teamID == "NY1") %>% 
  select(franchID)
```

**Which player hit the most home runs in the World Series?**

Here I want to use the data frame BattingPost that contains batting stats for all players who participated in playoff games.

I consider only World Series games and find the home run count for all players -- store in data frame S_HR.

```{r}
BattingPost %>% 
  filter(round == "WS") %>% 
  group_by(playerID) %>% 
  summarize(HR = sum(HR)) -> S_HR
head(S_HR)
```

Here are the leaders:

```{r}
S_HR %>% 
  arrange(desc(HR)) %>% 
  head()
```

Add names:

```{r}
inner_join(S_HR,
           select(People, 
                  playerID, nameFirst, nameLast),
           by = "playerID") %>% 
  mutate(Name = paste(nameFirst, nameLast)) %>% 
  select(Name, HR) %>% 
  arrange(desc(HR)) %>% 
  head(20)
```

